---
title: "Homework 6: Linear Models"
author: "Kristina Howell"
output: github_document
---

```{r settings, message = FALSE}
# These settings will be used throughout the document.

library(tidyverse)
library(modelr)
library(p8105.datasets)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis", 
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)

`%notin%` <- Negate(`%in%`)
```

## Problem 1

#### Data Tidying

```{r import_1, message=FALSE, warning=FALSE}
homicide_df = 
  
  # Import the data
  read_csv("./data/homicide-data.txt", na = c("", "NA", "Unknown")) %>% 
  
  # Create city_state variable and binary resolved variable
  mutate(
    city_state = str_c(city, state, sep = "_"),
    resolved = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest" ~ 0,
      disposition == "Closed by arrest" ~ 1)) %>% 
  
  # Remove unnecessary cities
  filter(city_state != "Tulsa_AL") %>% 
           # c("Tulsa_AL", "Dallas_TX", "Phoenix_AZ", "Kansas City_MO")) 
           
  # Include specific races
  filter(victim_race %in% c("White", "Black")) %>% 
  
  # Convert age to a numeric
  mutate(victim_age = as.numeric(victim_age),
         victim_race = as.factor(victim_race),
         victim_sex = as.factor(victim_sex)) %>% 
  
  # Retain only important variables
  select(city_state, resolved, victim_age, victim_race, victim_sex)
```


```{r}
# Create a dataframe with only Baltimore, MD
baltimore_df = 
  homicide_df %>% 
  filter(city_state == "Baltimore_MD")

# Fit a logistic regresison model 
glm(resolved ~ victim_age + victim_race + victim_sex,
    data = baltimore_df,
    family = binomial()) %>% 
  broom::tidy() %>% 
  
  # Obtain OR and CI
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)) %>% 
  
  # Retain necessary variables and create a table
  select(term, OR, starts_with("CI")) %>% 
  knitr::kable(digits = 3)
```

```{r}
models_results_df = 
  homicide_df %>% 
    nest(data = -city_state) %>% 
    mutate(
      models = 
        map(.x = data, 
            ~glm(resolved ~ victim_age + victim_race + victim_sex,
                 data = .x,
                 family = binomial())), 
      results = map(models, broom::tidy)) %>% 
    select(city_state, results) %>% 
    unnest(results) %>% 
    mutate(
      OR = exp(estimate),
      CI_lower = exp(estimate - 1.96 * std.error),
      CI_upper = exp(estimate + 1.96 * std.error)) %>% 
  select(city_state, term, OR, starts_with("CI"))
```

```{r}
models_results_df %>% 
  filter(term == "victim_sexMale") %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR)) +
    geom_point() +
    geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

## Problem 2

Import and clean data

```{r message=FALSE, warning=FALSE}
bwt_df = 
  
  # Import dataset using read_csv
  read_csv("./data/birthweight.csv") %>% 
  drop_na() %>% 
  
  # Retain only necessary variables
  select(bwt, blength, gaweeks, bhead, babysex, wtgain, mrace) %>% 
  
  # Relabel race categories and convert to factor
   mutate(
     mrace = case_when(
      mrace == 1 ~ "White",
      mrace == 2 ~ "Black",
      mrace == 3 ~ "Asian",
      mrace == 4 ~ "Puerto Rican",
      mrace == 8 ~ "Other"),
     mrace = as.factor(mrace),
    
  # Relabel baby's sex categories and convert to factor
    babysex = case_when(
      babysex == 1 ~ "Male",
      babysex == 2 ~ "Female"), 
   babysex = as.factor(babysex))

```

The first prespecified models are as follows:

* **Model 1:** birthweight ~ length at birth + gestational age

* **Model 2:** birthweight ~ head circumference | length | sex

My proposed model focuses on the mother's weight gain during pregnancy in pounds `wtgain` and mother's race `mrace` as well as the interaction between these two variables on the outcome variable, birthweight `btw`. This model is based on the underlying hypothesis of the effects of gestational diabetes mellitus (GDM) on birthweight, using mother's weight gain during pregnancy as a proxy measure. The risk of developing GDM is highest among Asian (particularly South Asian), Black, American Indian, and Hispanic women. Further, hese differences do not appear to be fully explained by differences in prepregnancy body mass index (BMI). More information is available [here](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4180530/). It is acknowledged that this hypothesis is a stretch for these variables, but it may be interesting to measure regardless. The model is as follows:

* **Model 3:** birthweight ~ wtgain + mrace + wtgain*mrace
















